---
title: Graphs-continuous variables
weight: 22
---
#### Learning Objectives

By the end of this section you should be able to:

* Understand how to plot continuous variables 
* Use boxplots, line plots , violin plots and 3D plots for continuous variables 


```{r, include=FALSE}
library(ggplot2)
load("EVS_UK.RData")
ggplot(EVS_UK) # this created an empty plot
EVS_UK$gender<- factor(EVS_UK$v225,
               levels = c(1,2),
               labels = c("Men", "Women"))

table(EVS_UK$gender)

EVS_UK$education <- ordered(EVS_UK$v243_r_weight, #here you specify that this is ordered variable
levels = c(1,2,3), # here you specify the values of the variable
labels = c("Low", "Medium", "High"))  #here you specify the names of the values

plot1<-ggplot(EVS_UK, aes(x = gender)) +
  geom_bar() +
  theme_classic()

plot1   # You can't view the plot unless you call it 

plot1<-plot1 + scale_x_discrete(breaks=c("1", "2"),
                      labels=c("Men", "Women"))
plot1
plot1<- plot1 +labs( y="Number of cases", x = "Responder's gender")
plot1<-plot1 + geom_bar(aes(fill = education))
plot1
plot1 + theme(legend.position="top")
plot1<-plot1 + theme(legend.position = c(0.25, 0.9),
          legend.direction = "horizontal") 
plot1 
```


### One continuous variable

For the following example we will use a variable describing citizens' satisfaction with the political system, the name of the variable is `v144`. We will use the `names()` function to give a meaningful name to the variable. 

The general formula is: 
names(my_data)[names(my_data) == "old.variable.name"] <- "new_variable_name"

```{r}
names(EVS_UK)[names(EVS_UK)=="v144"]<- "pol_sat"

table(EVS_UK$pol_sat)
```


```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot5 <- ggplot(na.omit(EVS_UK), aes(gender,pol_sat))

plot5<-plot5 + geom_boxplot(varwidth=T, fill="mediumpurple4") + 
    labs(title="Box plot", 
         subtitle="Satisfaction with the political system by gender",
         caption="Source: European Value Study (2019)",
         x="Gender",
         y="Satisfied with the political system")

plot5+theme_classic()

```

**See <a href="http://sape.inf.usi.ch/quick-reference/ggplot2/colour">here</a> for all other colours available** 


### Violin Plots

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot6 <- ggplot(na.omit(EVS_UK), aes(education, pol_sat))
plot6<-plot6 + geom_violin(fill="skyblue") + 
  labs(title="Violin plot", 
 subtitle="Satisfaction with the political system by level of education",
         caption="Source: European Value Study (2019)",
         x="Education",
         y="Satisfied with the political system")

plot6+theme_classic()
```


## Continuous variables

### Lines

In this section we will use the `EconomistData` dataset. 

```{r ,include=FALSE}
EconomistData <- read.csv("EconomistData.csv")
```

```{r}
plot7 <- ggplot(data = EconomistData, mapping = aes(x = HDI, y = CPI)) +
    geom_line() +
    facet_wrap(facets = vars(Region))+
  labs(title= "Lines",
        subtitle = "Human development index versus consumer price index",
         x = "Human Development Index",
         y = "Consumer Price Index") +
    theme_classic()

plot7
```

## 3D Plots: The plotly package

Let's try to plot some 3D plots. To do so we should install an additional package entitled `plotly`, `install.package("plotly")`.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(plotly)
```

```{r}
EconomistData <- read.csv("EconomistData.csv")
plot8<- plot_ly(data=EconomistData, x = ~HDI, y = ~CPI, z = ~HDI.Rank, color = ~Region, colors = c('#7B68EE', '#800000')) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'HDI'),
                     yaxis = list(title = 'CPI'),
                     zaxis = list(title = 'HDI Rank')))
plot8
```


Want to learn more about `plotly`? You may visit https://plot.ly/ggplot2/

# Recap


* The `plotly` package can be used to create 3D maps of continuous data.
* Continuous data can be represented using a variety of plot types. 
* `geom_boxplot()` is used to create box plots 
* `geom_violin()` is used to create violin plots
* `geom_line()` is used to create line plots

