---
title: Graphs-ordinal variables
weight: 22
---

#### Learning Objectives

By the end of this section you should be able to:

* Plot ordinal variables using ggplot2 
* Change or adjust the labels of a variable
* Use the density plot to represent ordinal variables.



```{r, include=FALSE}
library(ggplot2)
load("EVS_UK.RData")
ggplot(EVS_UK) # this created an empty plot
EVS_UK$gender<- factor(EVS_UK$v225,
               levels = c(1,2),
               labels = c("Men", "Women"))

table(EVS_UK$gender)

EVS_UK$education <- ordered(EVS_UK$v243_r_weight, #here you specify that this is ordered variable
levels = c(1,2,3), # here you specify the values of the variable
labels = c("Low", "Medium", "High"))  #here you specify the names of the values

plot1<-ggplot(EVS_UK, aes(x = gender)) +
  geom_bar() +
  theme_classic()

plot1   # You can't view the plot unless you call it 

plot1<-plot1 + scale_x_discrete(breaks=c("1", "2"),
                      labels=c("Men", "Women"))
plot1
plot1<- plot1 +labs( y="Number of cases", x = "Responder's gender")
plot1<-plot1 + geom_bar(aes(fill = education))
plot1
plot1 + theme(legend.position="top")
plot1<-plot1 + theme(legend.position = c(0.25, 0.9),
          legend.direction = "horizontal") 
plot1 
```

In this section we will use two variables describing anti-immigration attitudes, the first one is called `v184` and is asking respondents to agree or diagree with the following statement _"Immigrants take away jobs"_. The second variable is called `v186` and is asking respondents whether _"Immigrants are responsible for crime"_.

```{r}
names(EVS_UK)[names(EVS_UK) == "v184"] <- "immi_jobs"

table(EVS_UK$immi_jobs) # This is to check whether the function was correctly specified
```
We may add labels to the values using the functions we used at section 1:

{{% notice tip %}} If you only want to rename the name of the variable the easiest way is by using the `names()` function. 
{{% /notice %}}

```{r}
EVS_UK$immi_jobs <- ordered(EVS_UK$immi_jobs,#here you specify that this is ordered variable
levels = c(1,2,3,4,5), # here you specify the values of the variable
labels = c("Completely agree", "Agree", "Neither/nor","Disagree", "Completely disagree" ))  #here you specify the names of the values

table(EVS_UK$immi_jobs)
```

We will do the same for the variable describing whether "Immigrants tare responsible for crime". 

```{r}
EVS_UK$immi_crime <- ordered(EVS_UK$v186,#here you specify that this is ordered variable
levels = c(1,2,3,4,5),# here you specify the values of the variable
labels = c("Completely agree", "Agree", "Neither/nor","Disagree", "Completely disagree" ))  #here you specify the names of the values

table(EVS_UK$immi_jobs)
```




```{r}
plot3<-ggplot(na.omit(EVS_UK), aes(x = immi_crime),na.rm = TRUE) + geom_bar() + facet_grid(gender ~.)+
  theme_classic() #The na.omit() function will remove NAs

plot3
```

### Density Plots

```{r}
plot4 <- ggplot(na.omit(EVS_UK), aes(immi_crime))

plot4 + geom_density(aes(fill=factor(gender)), alpha=0.5) + 
    labs(title="Density plot", 
         subtitle="Anti-immigration attitudes by gender",
         caption="Source: European Value Study",
         x="Anti-immigration attitudes",
         fill="Gender")+
  theme_classic()
```

# Recap

* You can plot ordinal data using `geom_density()` in ggplot2.
* You can adjust or rename the ordinal data using the `ordered`,`levels` and `labels` functions in R.




