---
title: The model
weight: 27
---

```{r message=FALSE, include=FALSE}
library(dplyr)
load("EVS_UK_full.RData")
sub.sample<-EVS_UK_full %>% select(v102, v185, v189,v190, v191, v192, v193, v225, v226,v243_edulvlb,country)
sub.sample %>% filter(country == 826) 
sub.sample[sub.sample <=-1] <- NA
sub.sample[sub.sample <=-2] <- NA
sub.sample<-sub.sample %>% mutate(immi.jobs=(max(v185,na.rm=TRUE)-(v185)))
sub.sample<-sub.sample %>% mutate(born.country=(max(v189,na.rm=TRUE)-(v189)))
sub.sample<-sub.sample %>% mutate(respect.inst=(max(v190,na.rm=TRUE)-(v190)))
sub.sample<-sub.sample %>% mutate(country.ancestry=(max(v191,na.rm=TRUE)-(v191)))
sub.sample<-sub.sample %>% mutate(speak.lang=(max(v192,na.rm=TRUE)-(v192)))
sub.sample<-sub.sample %>% mutate(share.cultr=(max(v193,na.rm=TRUE)-(v193)))
sub.sample<-sub.sample %>% mutate(age=2017-v226)
table(sub.sample$age)
sub.sample<-sub.sample %>% rename(education=v243_edulvlb)
sub.sample<-sub.sample %>% rename(gender=v225)
sub.sample<-sub.sample %>% rename(left_right=v102)
```




We are now ready to conduct our analysis. We will use the `lm()` function.

```{r mregression, message=FALSE, warning=FALSE}
library(lme4)
model.1<-lm(immi.jobs~ born.country+respect.inst+country.ancestry+speak.lang+share.cultr+left_right+education+gender+age,data=sub.sample)
summary(model.1)
```

```{r arm, message=FALSE, warning=FALSE}
library(dotwhisker)
```

```{r}
dwplot(list(model.1), vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) %>% # plot line at zero _behind_ coefs
    relabel_predictors(c(immi.jobs = "Immigrants take jobs",                       
                         born.country = "Born in country", 
                         respect.inst = "Respect institutions", 
                         country.ancestry = "Country ancestry", 
                         speak.lang = "Speak language", 
                         share.cultr = "Share culture",
                         left_right="Left-Right",
                         education="Education",
                         gender="Gender",
                         age="Age")) +
     theme_classic() + xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Predicting anti-immigration attitudes") +
     theme(plot.title = element_text(face="bold"),
           legend.position = c(0.01, 0.03),
           legend.justification = c(0, 0), 
           legend.background = element_rect(colour="grey80"),
           legend.title = element_blank()) 

```

From our analysis we can see that three of our variables are not statistically significant, namely age, respect towards institutions, and left-right. Let's construct another model but this time we will exclude the three variables that are not statistically significant. 


```{r}
model.2<-lm(immi.jobs~ born.country+country.ancestry+speak.lang+share.cultr+education+gender,data=sub.sample)
summary(model.2)
```

We can now plot both models and compare the results.

```{r 2plots}
dwplot(list(model.1, model.2), vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) %>% # plot line at zero _behind_ coefs
    relabel_predictors(c(immi.jobs = "Immigrants take jobs",                       
                         born.country = "Born in country", 
                         respect.inst = "Respect institutions", 
                         country.ancestry = "Country ancestry", 
                         speak.lang = "Speak language", 
                         share.cultr = "Share culture",
                         left_right="Left-Right",
                         education="Education",
                         gender="Gender",
                         age="Age")) +
     theme_classic() + xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Predicting anti-immigration attitudes") +
     theme(plot.title = element_text(face="bold"),
           legend.position = c(0.01, 0.03),
           legend.justification = c(0, 0), 
           legend.background = element_rect(colour="grey80"),
           legend.title = element_blank()) 
```