---
title: The ggplot package
weight: 21
---
#### Learning Objectives

By the end of this section you will:

* Learn the basic concepts associated the grammar of graphics such as aesthetic mapping.
* Be introduced to some initial steps to preparing your data for plotting graphs.


## The ggplot package

The most common data visualisation package in RStudio is ggplot. We may install ggplot by using the `install.packages()` function. We write `install.packages("ggplot2")` and we call the package by using the `library()` function.

```{r libraryggplot}
library(ggplot2)
```

First of all, as most packages in RStudio the ggplot team created a cheatsheet. You may find it here
[ggplot Cheatsheet](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf)

If we call the `ggplot()` function then we will create an empty canvas. 
We start by loading our dataset entitled **EVS_UK.RData**

```{r emptyplot}
load("EVS_UK.RData")
ggplot(EVS_UK) # this created an empty plot
```

the next step is to specify the variables we would like to use, as you know we cannot plot the whole dataset! 

To specify which variables we would like to plot we have to include in the function the so called **aes()** section that specifies the **aesthetic mappings**, in other words, this section specifies how to map our variables. 


Let's start by creating a bar to see how the **aes** section works - you already know how to do that but this time we will give a name to our plot.


In our analysis we will use two socio-demographics variables gender and education. Gender is a dichotomous variable describing gender. Education is an ordinal variable with three levels - low, medium, and high.

In the dataset the code of the variable describing gender is `EVS_UK$v225`, and the name of the variable describing education is `EVS_UK$v243_r_weight`. 

Before we proceed, we will give meaningful names to our variable swhile at the same time we will make sure that our new variables are categorical (_factor_) variable for gender- as it should be. We can easily do that by using the assignment operator $<-$ and the `as.factor` function.  


```{r variable name}
EVS_UK$gender<- factor(EVS_UK$v225,
               levels = c(1,2),
               labels = c("Men", "Women"))

table(EVS_UK$gender)
```
**Note**: at the left side of the equation I specified the dataset at which my new variable belongs to - that is `EVS_UK`. 

We will do the same for the variable describing education and is currently named `v243_r_weight`. 

Recall, Lecture 2 (Descriptive Statistics) during the lab session we used the same function to rename `v243_r_weight`to `education`.

In this example education is an ordinal variable with three levels- low, medium, high. 

```{r}
EVS_UK$education <- ordered(EVS_UK$v243_r_weight, #here you specify that this is ordered variable
levels = c(1,2,3), # here you specify the values of the variable
labels = c("Low", "Medium", "High"))  #here you specify the names of the values
```

# Recap

* ggplot2 refers to the grammar of graphic and is the main package in the 'tidyverse' for plotting data.
* Aesthetic mapping or `aes()` is the function used tell ggplot2 which variables to use in our plot.
* Before we plot our data we need to ensure that our labels are meaningful.


