---
title: Predicted probabilities
weight: 32
---
```{r, include=FALSE}
load("bes.RData")
library(dplyr)
eco_voting<-bes %>%select(b02,Age,edlevel,e01,l01,l02,l03,l04,p01)
View(eco_voting)
eco_voting[eco_voting <=-1] <- NA
eco_voting[eco_voting <=-2] <- NA
eco_voting<-eco_voting %>% mutate(incumbent=case_when(b02==2~1, b02==1~0, b02>=3~0)) 
eco_voting<-eco_voting %>% rename(left_right=e01, egocentric_retro=l01,egocentric_pros=l02,sociotropic_pros=l03,sociotropic_retro=l04,brexit=p01)
table(eco_voting$b02)
table(eco_voting$incumbent)
logit.model<-glm(incumbent~ sociotropic_pros+egocentric_retro+left_right, data=eco_voting,family = binomial(link = "logit"))
summary(logit.model)
```

## Predicted probabilities

As you may have already realised, log-odds are not straight-forward. This is why we use the `predict()` function which give us predictions for Y, the dependent variable. 

```{r}
probabilities<-predict(logit.model,type = "response")
```


## The effects package

The final section include's the use of the `effects` package developed by Fox, et al. The `effects` package creates plots for various statistical models. In this example I will show you how it works with a logistic model- but it works with linear models, mixed effects models and many other.  

First you have to install the `effects` package. The easiest way to visualise your model is by using the `allEffects()` function in combination with the `plot()` function. 

```{r message=FALSE, warning=FALSE}
library(effects)
plot(allEffects(logit.model))
```

**References**

Fox, J. (2003) Effect displays in R for generalised linear models. Journal of Statistical Software 8:15, 1–27, http://www.jstatsoft.org/v08/i15/.


